### 2017-09-13 manuscript.Rmd
updates to manuscript, added new sections

---
layout: review, 11pt
linenumbers: true
title: "rOpenSci tools for textmining open source science literature"
author:
  - name: Scott Chamberlain
    affiliation: cstar
    email: scott(at)ropensci.org
    footnote: Corresponding author
address:
  - code: cstar
    address: |
      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA
abstract: |
      Corresponding Author:

      Scott Chamberlain

      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA

      Email address: [scott@ropensci.org](mailto:scott@ropensci.org)

      \newpage

      Background. xxxx.

      Methods. xxxx.

      Results. xxxx.

      Discussion. xxxx.

bibliography: components/references.bib
csl: components/peerj.csl
documentclass: components/elsarticle

output:
  pdf_document:
    template: components/elsarticle.latex
    keep_tex: true
    fig_caption: true
---

```{r compile-settings, include=FALSE}
library("methods")
library("knitr")
opts_chunk$set(
  tidy = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = 1,
  comment = '#>',
  collapse = TRUE,
  verbose = TRUE
)

basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile'))
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))

# tibble options
options(tibble.max_extra_cols = 10)
```

\newpage

# Introduction

There's likely more than 100 million articles published
(source: Crossref API), representing an enormous amount
of knowledge.  In addition to simply reading these articles,
they contain a vast trove of information of interest to
researchers.

For example, many researchers are interested in statistical
outcomes of articles: questions about P-values, about effect
sizes, and more. With regard to effect sizes, these are
of particular interest, as they are often combined in
meta-analyses to draw broad conclusions about a particular
question.

Text-mining is the broad term associated with pulling information
out of articles. Given the importance of text-mining, good
text-mining tools are needed to make it easier for researchers
to do.

Here, we do an overview of text-mining tools in the R programming
language. We do not cover analysis tools per se, but rather those
tools for searching for, acquiring, and "mashing up" text.

# Digital articles: technical aspects

Of digital articles some of which are available digitally,
and some of which are not. Those that are digital can be split
into two groups: easily machine readable and non-machine readable.

The machine readable articles are those in XML, JSON, or plain text
format. The former two, XML and JSON, are ideal of the machine
readable types because they are structured data, whereas
plain text has no structure - it's simply a long set of characters
with line breaks and spaces in between.

Of the non-machine readable kind, there's PDFs. These can be broken out
into two groups: text based PDFs and scanned PDFs. The former are
converted from digital versions of various kinds (MS Word, OpenOffice,
markdown, etc.), while the latter are PDFs created by scanning
in print articles for which there is no digital version.

# Digital articles: the access landscape

Acces to full-text is the holy grail in text-mining. Some use cases can
get by with article metadata (authors, title, etc.), some with abstracts,
but many use cases need full-text.

The landscape of access to full-text is a extremely hetergeous, with the
majority of variation along the publisher axis. The major hurdle are
paywalls. The majority of articles are published by the big three
publishers - Wiley, Springer, Elsever - and the majority of their articles
are behind paywalls.

A promising sign is that there's an increasing number of open access
publishers. xxxx.

# The discovery problem (maybe remove section)

xxx

# Data sources

There is increasing open source scientific literature content available
online. However, only a small proportion of scientific journals provide
access to their full content; whereas, most publishers provide open
access to their metadata only (most often through Crossref; Table 1).

Table 1. Sources of scientific literature, their content type provided via web services,
whether rOpenSci has an R packages for the service, and where to find the API documentation.

Data Provider                      | Content Type            | rOpenSci Pkg?            | API Documentation
---------------------------------- | ---------------------   | ------------------------ | -----------------
Crossref                           | Metadata only           | rcrossref                | [^1]
DataCite                           | Metadata only           | rdatacite                | [^2]
Biodiversity Heritage Library      | Full content/Metadata   | rbhl                     | [^3]
Public Library of Science (PLoS)   | Full text/altmetrics    | rplos                    | [^4]
Scopus (Elsevier)                  | Full content/Metadata   | fulltext                 | [^5]
arXiv                              | Full content/Metadata   | aRxiv                    | [^6]
Biomed Central (via Springer)      | Full content/Metadata   | fulltext                 | [^7]
bioRxiv                            | Full content/Metadata   | fulltext                 | [^8]
PMC/Pubmed (via Entrez)            | Full content/Metadata   | rentrez                  | [^9]
Microsoft Academic Search          | Metadata                | fulltext/microdemic      | [^10]

[^1]: http://api.crossref.org
[^2]: https://support.datacite.org/docs/api
[^3]: http://bit.ly/KYQ1Rd
[^4]: http://api.plos.org/solr
[^5]: http://bit.ly/J9S616
[^6]: https://arxiv.org/help/api/index
[^7]: https://dev.springer.com/
[^8]: http://www.biorxiv.org/
[^9]: https://www.ncbi.nlm.nih.gov/books/NBK25500
[^10]: https://azure.microsoft.com/en-us/services/cognitive-services

The following is a synopsis of the major data sources and associated R tools.

## Crossref

Crossref is a non-profit that creates (or "mints") Digital Object Identifiers (DOIs).
In addition, they maintain metadata associated with each DOI. The metadata ranges
from simple (including author, title, dates, DOI, type, publisher) to including
number of citations to the article, as well as references in the article, and even
abstracts.

Crossref does have a text-mining opt-in program for publishers. The result of this is that some
publishers deposit URLs for full text content of their articles. The majority of these links
are pay-walled, while some are open access. Using any of the various tools for working with
Crossref data, you can filter your search to get only articles with full text links, and further
to get only articles with full text links that are open access.

The main interface for Crossref in R is [rcrossref][]. Parallel interfaces are available in
Ruby ([serrano][]) and Python ([habanero][]).

## Pubmed

Pubmed is a corpus/website of NIH funded research ...

# How to text mine from R: Three case studies

### Case study 1

```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
```

### Case study 2

```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
```

### Case study 3

```{r city-diversity, child="use-cases/city-diversity.Rmd", eval=FALSE}
```

# Conclusions and future directions

xxxx

# Acknowledgments

xxxx

# Data Accessibility

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
<https://github.com/ropensci/textmine>. Software can be found at
<https://github.com/ropensci/xxx>, xxxx, all under MIT licenses.

# References

[rcrossref]: https://github.com/ropensci/rcrossref
[serrano]: https://github.com/sckott/serrano
[habanero]: https://github.com/sckott/habanero