### 2019-04-05 manuscript.Rmd
use cases filled out a bit, updates to data provider sections

---
layout: review, 11pt
linenumbers: true
title: "rOpenSci tools for accessing science literature for textmining"
author:
  - name: Scott Chamberlain
    affiliation: cstar
    email: myrmecocystus(at)gmail.com
    footnote: Corresponding author
address:
  - code: cstar
    address: |
      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA
abstract: |
      Corresponding Author:

      Scott Chamberlain

      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA

      Email address: [myrmecocystus@gmail.com](mailto:myrmecocystus@gmail.com)

      \newpage

      Background. xxxx.

      Methods. xxxx.

      Results. xxxx.

      Discussion. xxxx.

bibliography: components/references.bib
csl: components/peerj.csl
documentclass: components/elsarticle

output:
  pdf_document:
    template: components/elsarticle.latex
    keep_tex: true
    fig_caption: true
---

```{r compile-settings, include=FALSE}
library("methods")
library("knitr")
opts_chunk$set(
  tidy = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = 1,
  comment = '#>',
  collapse = TRUE,
  verbose = TRUE
)

basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile'))
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))

# tibble options
options(tibble.max_extra_cols = 10)
```

\newpage

# Introduction

There's more than 100 million articles published
(source: Crossref API), representing an enormous amount
of knowledge.  In addition to simply reading these articles,
they contain a vast trove of information of interest to
researchers for machine aided questions.

For example, many researchers are interested in statistical
outcomes of articles: questions about P-values, about effect
sizes, and more. With regard to effect sizes, these are
of particular interest, as they are often combined in
meta-analyses to draw broad conclusions about a particular
question.

Text-mining is the broad term associated with pulling information
out of articles. Given the importance of text-mining, good
text-mining tools are needed to make it easier for researchers.
In particular, the R programming language is used widely throughout
many academic fields and thus tools in R for text mining are of
particular importance.

Here, we present an overview of text-mining tools in the R programming
language. We do not cover analysis tools per se, but rather those
tools for searching for, acquiring, and "mashing up" text.

# Digital articles: technical aspects

Those articles that are digital can be split into two groups:
easily machine readable and non-machine readable.

The machine readable articles are those in XML, JSON, or plain text
format. The former two, XML and JSON, are ideal for the machine
readable types because they are structured data, whereas
plain text has no structure - it's simply a set of characters
with line breaks and spaces in between.

Of the non-machine readable kind, there's PDFs. These can be broken out
into two groups: text based PDFs and scanned PDFs. The former are
converted from digital versions of various kinds (MS Word, OpenOffice,
markdown, etc.), while the latter are PDFs created by scanning
in print articles for which there is no digital version.

# Digital articles: the access landscape

Acces to full-text is the holy grail in text-mining. Some use cases can
get by with article metadata (authors, title, etc.), some with abstracts,
but many use cases need full-text.

The landscape of access to full-text is extremely hetergeous, with the
majority of variation along the publisher axis. The major hurdle is
paywalls. The majority of articles are published by the big three
publishers - Wiley, Springer, Elsevier - and the majority of their articles
are behind paywalls.

A promising sign is an increasing number of open access publishers, yet
these represent a very small portion of the total articles (XXXXX) (ref.).

With respect to paywalled articles, access varies by institution, depending
on what each institution decided to pay for. In addition, some users may
not realize access varies with IP address so that access from campus vs.
from home (if not on a VPN) will drastically differ.

We can not end this section without mentioning SciHub. This is a last
resort option for many probably, providing access to full text of articles
that are normally paywalled. No tools in this manuscript provide access
to SciHub.

# The discovery problem

xxx

#

xxx

# Data sources

There is increasing open access scientific literature content available
online. However, only a small proportion of scientific journals provide
access to their full content; whereas, most publishers provide open
access to their metadata only (most often through Crossref; Table 1).
The following is a synopsis of the major data sources and associated R
tools.

\newpage
Table 1. Sources of scientific literature, their content type provided via
web services, whether rOpenSci has an R packages for the service, and
where to find the API documentation.

Data Provider                      | Content Type            | rOpenSci Package         | Documentation
---------------------------------- | ---------------------   | ------------------------ | -----------------
Crossref                           | Metadata only           | rcrossref/crminer        | [^1]
DataCite                           | Metadata only           | rdatacite                | [^2]
Biodiversity Heritage Library      | Full content/Metadata   | rbhl                     | [^3]
Public Library of Science (PLoS)   | Full text/altmetrics    | rplos                    | [^4]
Scopus (Elsevier)                  | Full content/Metadata   | fulltext                 | [^5]
arXiv                              | Full content/Metadata   | aRxiv                    | [^6]
Biomed Central (via Springer)      | Full content/Metadata   | fulltext                 | [^7]
bioRxiv                            | Full content/Metadata   | fulltext                 | [^8]
PMC/Pubmed (via Entrez)            | Full content/Metadata   | rentrez                  | [^9]
Europe PMC                         | Full content/Metadata   | europepmc                | [^10]
Microsoft Academic Search          | Metadata                | fulltext/microdemic      | [^11]
Directory of Open Access Journals  | Metadata                | jaod                     | [^12]
JSTOR Data for Research            | Full content            | jstor                    | [^13]
ORCID                              | Metadata                | rorcid                   | [^14]
Wikimedia's Citoid                 | Citations               | rcitoid                  | [^15]
Open Citation Corpus               | Citations               | citecorp                 | [^16]

[^1]: https://api.crossref.org
[^2]: https://support.datacite.org/docs/api
[^3]: http://bit.ly/KYQ1Rd
[^4]: http://api.plos.org/solr
[^5]: http://bit.ly/J9S616
[^6]: https://arxiv.org/help/api/index
[^7]: https://dev.springer.com/
[^8]: http://www.biorxiv.org/
[^9]: https://www.ncbi.nlm.nih.gov/books/NBK25500
[^10]: https://azure.microsoft.com/en-us/services/cognitive-services
[^11]: https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97
[^12]: https://doaj.org/api/v1/docs
[^13]: https://www.jstor.org/dfr/
[^14]: https://pub.orcid.org/
[^15]: https://en.wikipedia.org/api/rest_v1/#/Citation/getCitation
[^16]: http://opencitations.net/

## Crossref/Datacite

Crossref is a non-profit that creates (or "mints") Digital Object Identifiers (DOIs).
In addition, they maintain metadata associated with each DOI. The metadata ranges
from simple (including author, title, dates, DOI, type, publisher) to including
number of citations to the article, as well as references in the article, and even
abstracts. At the time of writing they hold 100 million DOIs.

One can search by DOI or search citation data to get citations. In addition, Crossref has a
text-mining opt-in program for publishers. The result of this is that some
publishers provide URLs for full text content of their articles. The majority of these links
are pay-walled, while some are open access. Using any of the various tools for working with
Crossref data, you can filter your search to get only articles with full text links, and further
to get only articles with full text links that are open access.

The main interfaces for Crossref in R are [rcrossref][] and [crminer][]. Similar interfaces
are available in Ruby ([serrano][]) and Python ([habanero][]).

Datacite is similar to Crossref, but focuses on datasets instead of articles. The main
interface for Datacite in R is [rdatacite][].

## Biodiversity Heritage Library

The Biodiversity Heritage Library (BHL) houses scans of biodiversity books, and provides
web interfaces and APIs to query and fetch those data. They also provide text of the scanned
pages. The main R interace to BHL is through [rbhl][].

## Public Library of Science

The Public Library of Science (PLOS) is one of the largest open access only publishers. They
as of this writing have published 2.1 million articles. One of the strongs advantages of PLOS
is that they provide an API to their Solr instance, which is a very flexible way to
search their articles. The main R interace to PLOS is through [rplos][].

## Elsevier/Scopus

Elsevier is one of the largest publishers. Most of their articles are not open access. However,
they have a numbrer of advantages if you have access to their articles: they are one of the few
publishers to provide machine readable XML (many publishers do have XML versions of articles,
but do not provide it); they are one of the few (two) publishers part of Crossref's text
and data mining program. The packages [fulltext][] and [crminer][] can be used to access
Elsevier articles through Crossref's TDM program. There's an interface to Scopus article
search within [fulltext][].

## arXiv/bioRxiv

arXiv and bioRxiv are preprint publishers, the former in existence for many years, and the
latter new on the scene. You can access articles from these publishers through [fulltext][].
arXiv does provide a web API that we hook into; bioRxiv does not, but we can get you
articles nonetheless.

## Pubmed/PMC/Europe PMC

Pubmed/PMC is a corpus/website of NIH funded research in the United States; while Europe PMC
is an equivalent for the European Union. You can access articles from Pubmed/PMC through
[fulltext][], and for Europe PMC through [europepmc][].

## Microsoft Academic Research

Microsoft Academic Research (MAR) is a search engine for research articles. You can use their
GUI web interface to search, and they provide APIs for programmatic access. The R interface
for MAR is [microdemic][]; and [fulltext][] hooks into `microdemic` as well for article search
and abstract retrieval.

## Directory of Open Access Journals

xxxxx

## JSTOR

xxxxx

## ORCID

xxxxx

## Citoid/Open Citation Corpus

xxx

# How to text mine from R: Three case studies

### Case study 1: Citation mining

```{r citation-mining, child="use-cases/citation-mining.Rmd", eval=TRUE}
```

### Case study 2: Abstract mining

```{r abstract-mining, child="use-cases/abstract-mining.Rmd", eval=TRUE}
```

### Case study 3: Full text mining

```{r full-text-mining, child="use-cases/full-text-mining.Rmd", eval=TRUE}
```

# Future directions

xxxx

# Acknowledgments

xxxx

# Data Accessibility

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
<https://github.com/ropensci/textmine>. Software can be found at
<https://github.com/ropensci/xxx>, xxxx, all under MIT licenses.

# References

[rcrossref]: https://github.com/ropensci/rcrossref
[rorcid]: https://github.com/ropensci/rorcid
[fulltext]: https://github.com/ropensci/fulltext
[microdemic]: https://github.com/ropensci/microdemic
[arXiv]: https://github.com/ropensci/arXiv
[rdatacite]: https://github.com/ropensci/rdatacite
[rbhl]: https://github.com/ropensci/rbhl
[rentrez]: https://github.com/ropensci/rentrez
[rplos]: https://github.com/ropensci/rplos
[crminer]: https://github.com/ropensci/crminer
[europepmc]: https://github.com/ropensci/europepmc
[jaod]: https://github.com/ropensci/jaod
[jstor]: https://github.com/ropensci/jstor
[rcitoid]: https://github.com/ropenscilabs/rcitoid
[citecorp]: https://github.com/ropenscilabs/citecorp
[serrano]: https://github.com/sckott/serrano
[habanero]: https://github.com/sckott/habanero
[quanteda]: https://quanteda.io/
[pubchunks]: https://github.com/ropensci/pubchunks