### 2019-04-05 use-cases/abstract-mining.Rmd

```{r include=FALSE}
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = '#>',
  collapse = TRUE,
  cache.path = "cache/"
)
```

Sometimes you just need abstracts for your research question. The benefit of 
only needing abstracts, and not need full text, is that there's many more
articles that will have abstracts available than have their full text 
available.

As an example, let's say you xxxx

```{r}
library("fulltext")
```

_xxxxx_

Using `fulltext`:

```{r}
res <- ft_search("ecology", from = "crossref",
  crossrefopts = list(filter = c(has_abstract = TRUE)))
ids <- res$crossref$data$doi
out <- ft_abstract(x = ids, from = "crossref")
abstracts <- vapply(out$crossref, "[[", "", "abstract")
```

Using [quanteda][], read the abstracts into a corpus

```{r}
library("quanteda")
corp <- corpus(abstracts)
docvars(corp) <- ids
```

Get a summary of the abstracts

```{r}
summary(corp)
```

Use the `kwic()` function to see a word in context across
the abstracts

```{r}
kwic(corp, pattern = "ecology")
```