### 2019-04-12 manuscript.Rmd
work on ms

---
layout: review, 11pt
linenumbers: true
title: "R tools for accessing research literature for text mining"
author:
  - name: Scott Chamberlain
    affiliation: cstar
    email: myrmecocystus(at)gmail.com
    footnote: Corresponding author
address:
  - code: cstar
    address: |
      rOpenSci, Museum of Paleontology, University of California, Berkeley, CA, USA
abstract: |
  Text-mining is a powerful method for answering research questions. However,
  getting texts to extract information can be a daunting and complicated task.
  The primary reason for this is publishers. There are thousands of different
  publishers, each with their own licenses, URL patterns, access options, and
  more. Layered on top of that is the varied access each user has based on
  their institutional affiliation. Here, I introduce a suite of software packages
  in the R programming language for fetching texts. The tapestry of different
  publishers, access levels, and other factors requires a patchwork of approaches
  for getting texts to users. The flagship R package called fulltext attempts
  to simplify search and retrieval of texts for text-mining. The fulltext package,
  along with many others, make acquiring texts easier than ever, facilitating
  answering research questions with text-mining.
bibliography: components/references.bib
csl: components/peerj.csl
documentclass: components/elsarticle
output:
  pdf_document:
    template: components/elsarticle.latex
    keep_tex: true
    fig_caption: true
---

```{r compile-settings, include=FALSE}
library("methods")
library("knitr")
opts_chunk$set(
  tidy = FALSE,
  warning = FALSE,
  message = FALSE,
  cache = 1,
  comment = '#>',
  collapse = TRUE,
  verbose = TRUE
)

basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile'))
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))

# tibble options
options(tibble.max_extra_cols = 10)
```

\newpage

# Introduction

There's more than 100 million research articles published
(Crossref API: <https://github.com/CrossRef/rest-api-doc>), representing
an enormous amount of knowledge.  In addition to simply reading these
articles, the articles contain a vast trove of information of interest to
researchers for machine aided questions [@Kong2018; @Usai2018].

For example, many researchers are interested in statistical
outcomes of articles: questions about P-values, about effect
sizes, and more. With regard to effect sizes, these are
of particular interest, as they are often combined in
meta-analyses to draw broad conclusions about a particular
question.

Text-mining is the broad term associated with pulling information
out of articles. Given the importance of text-mining, good
text-mining tools are needed to make it easier for researchers to do.
Graphical user interface (GUI) based text mining tools are available
[e.g., @ba2016interoperability] and some research papers have used them
[@Chaix2019], but given the urgent recent call to action for more reproducible
research [@OSC2015; @Camerer2016; @Camerer2018],
we must move away from GUI based tools as fast as possible.

In particular, the R programming language is used widely throughout many
academic fields and thus tools in R for text mining are of particular
importance because they can be adopted by academics rapidly.

Here, I present an overview of text-mining tools in the R programming
language, not for text-mining analysis, but rather those
tools for searching for, acquiring, and extracting particular chunks
of texts (e.g., title, abstract, authors). Most of the packages are
part of the rOpenSci suite (https://ropensci.org/).

# Digital articles: technical aspects

Those articles that are digital (which in theory includes all articles)
can be split into two groups: machine readable and non-machine
readable.

The machine readable articles are those in XML[^xml], JSON[^json], or
plain text format. The former two, XML and JSON, are the best machine
readable types because they are structured data[^sd], whereas
plain text has no structure - it's simply a set of characters
with line breaks and spaces in between.

Of the non-machine readable types, the most noteable is the Portable
Document Format (PDF)[^pdf]. These can be broken out into two groups: text based
PDFs and scanned PDFs. The former are converted from digital versions of
various kinds (MS Word, OpenOffice, LaTeX, markdown, etc.), while the
latter are created by scanning print articles to a PDF format. Text-based
PDFs are much better for text-mining purposes as plain text can be extracted
easily in R with [pdftools][], a binding to [libpoppler][]. However, with
scanned PDFs, text must be extracted using Optimal Character Recognition
(OCR; see R package [tesseract][]), which isn't always a clean solution,
especially compared to true text based PDFs.

The reality in scholarly publishing is all publishers, if they provide
any access to their articles, only provide PDF format. Very few publishers,
with some quite large (Elsevier, Pensoft, PLOS), provide XML format. Although
most publishers most likely have the XML behind each of their articles,
they for some indefensible reason do not share it - making text-mining more
difficult. Some provide plain text (Elsevier). I only know of one publisher
that provides full text as JSON (PLOS). Thus, text-mining, in most cases,
will require extracting text from PDFs.

[^xml]: https://www.w3.org/TR/xml/
[^json]: https://tools.ietf.org/html/rfc7159
[^sd]: https://en.wikipedia.org/wiki/Data_model
[^pdf]: https://en.wikipedia.org/wiki/PDF

# Digital articles: the access landscape

Acces to full-text is the holy grail in text-mining. Some use cases can
get by with article metadata (authors, title, etc.), some with abstracts,
but many use cases require full-text.

The landscape of access to full-text is extremely hetergeous, with the
majority of variation along the publisher axis. The major hurdle is
paywalls. The majority of articles are published by the big three
publishers - Wiley, Springer, Elsevier - and the majority of their articles
are behind paywalls.

A promising sign is an increasing number of open access articles, yet
open access articles represent a small percent of all articles: an estimate in
2018 said that 28% of the scholarly literature was open access [@Piwowar2018].

With respect to paywalled articles, access varies by institution, depending
on each institution's publisher contracts. MORE ABOUT THIS ...

Some may not realize access to articles varies with IP address so that
access from campus vs. from home (if not on a VPN) will drastically differ.
Sometimes a VPN is required, and this can provide a significant technical
hurdle to users attempting to do text-mining work.

One final hurdle in text-mining comes unsurprisingly from Elsevier. They
use so-called "fences" for programmatic access. That is, even if a person
trying to get an article programmatically their institution has access to
and they have access to, and they are on the correct IP address, they may
still not get access to an Elsevier article. Elsevier puts in place these
fences and only if you contact their technical team directly can you
get these fences removed, and only then on a per institution basis.

I can not end this section without mentioning SciHub. This is a last
resort option for many probably (or possibly first, depending on your level
of access), providing access to full text of articles that are normally
paywalled. No tools in this manuscript provide access to SciHub.

# The discovery problem

A text-mining project starts with a question. From that question,
researchers then attempt to acquire scholarly articles for
text-mining. Finding those articles is not altogether straight-forward.

There are many places to search for articles; a non-exhaustive list:
Google Scholar, Microsoft Academic Research, Scopus, ScienceDirect,
Web of Science, Pubmed/Entrez, Europe PMC, Directory of Open Access Journals,
Open Knowledge Maps, and more. It's probably difficult to know where the
best place is to search. Some of these are paywalled (e.g., Web of Science),
and some are not.

The most important aspect about any source for article search with respect to
reproducible research is being able to use the data source programmatically.
Of those listed above, the following can  be used programmatically:
Microsoft Academic Research, Scopus, ScienceDirect, Pubmed/Entrez, Europe PMC,
and Directory of Open Access Journals. All of these are included in the R
package [fulltext][], discussed further below.

# Data sources

There is increasing open access scientific literature content available
online. However, only a small proportion of scientific journals provide
access to their full content; whereas, most publishers provide open
access to their metadata only (most often through Crossref; Table 1).
The following is a synopsis of the major data sources and associated R
tools.

\newpage
Table 1. Sources of scientific literature, their content type provided via
web services, whether rOpenSci has an R packages for the service, and
where to find the API documentation.

Data Provider                      | Content Type            | rOpenSci Package         | Documentation
---------------------------------- | ---------------------   | ------------------------ | -----------------
Crossref                           | Metadata only           | rcrossref/crminer        | [^1]
DataCite                           | Metadata only           | rdatacite                | [^2]
Biodiversity Heritage Library      | Full content/Metadata   | rbhl                     | [^3]
Public Library of Science (PLoS)   | Full text/altmetrics    | rplos                    | [^4]
Scopus (Elsevier)                  | Full content/Metadata   | fulltext                 | [^5]
arXiv                              | Full content/Metadata   | aRxiv                    | [^6]
Biomed Central (via Springer)      | Full content/Metadata   | fulltext                 | [^7]
bioRxiv                            | Full content/Metadata   | fulltext                 | [^8]
PMC/Pubmed (via Entrez)            | Full content/Metadata   | rentrez                  | [^9]
Europe PMC                         | Full content/Metadata   | europepmc                | [^10]
Microsoft Academic Search          | Metadata                | fulltext/microdemic      | [^11]
Directory of Open Access Journals  | Metadata                | jaod                     | [^12]
JSTOR Data for Research            | Full content            | jstor                    | [^13]
ORCID                              | Metadata                | rorcid                   | [^14]
Wikimedia's Citoid                 | Citations               | rcitoid                  | [^15]
Open Citation Corpus               | Citations               | citecorp                 | [^16]

[^1]: https://api.crossref.org
[^2]: https://support.datacite.org/docs/api
[^3]: http://bit.ly/KYQ1Rd
[^4]: http://api.plos.org/solr
[^5]: http://bit.ly/J9S616
[^6]: https://arxiv.org/help/api/index
[^7]: https://dev.springer.com/
[^8]: http://www.biorxiv.org/
[^9]: https://www.ncbi.nlm.nih.gov/books/NBK25500
[^10]: https://azure.microsoft.com/en-us/services/cognitive-services
[^11]: https://dev.labs.cognitive.microsoft.com/docs/services/56332331778daf02acc0a50b/operations/565d9001ca73072048922d97
[^12]: https://doaj.org/api/v1/docs
[^13]: https://www.jstor.org/dfr/
[^14]: https://pub.orcid.org/
[^15]: https://en.wikipedia.org/api/rest_v1/#/Citation/getCitation
[^16]: http://opencitations.net/

## Crossref/Datacite

Crossref is a non-profit that creates (or "mints") Digital Object Identifiers (DOIs).
In addition, they maintain metadata associated with each DOI. The metadata ranges
from simple (including author, title, dates, DOI, type, publisher) to including
number of citations to the article, as well as references in the article, and even
abstracts. At the time of writing they hold 100 million DOIs.

One can search by DOI or search citation data to get citations. In addition, Crossref has a
text-mining opt-in program for publishers. The result of this is that some
publishers provide URLs for full text content of their articles. The majority of these links
are pay-walled, while some are open access. Using any of the various tools for working with
Crossref data, you can filter your search to get only articles with full text links, and further
to get only articles with full text links that are open access.

The main interfaces for Crossref in R are [rcrossref][] and [crminer][]. Similar interfaces
are available in Ruby ([serrano][]) and Python ([habanero][]).

Datacite is similar to Crossref, but focuses on datasets instead of articles. The main
interface for Datacite in R is [rdatacite][].

## Biodiversity Heritage Library

The Biodiversity Heritage Library (BHL) houses scans of biodiversity books, and provides
web interfaces and APIs to query and fetch those data. They also provide text of the scanned
pages. The main R interace to BHL is through [rbhl][].

## Public Library of Science

The Public Library of Science (PLOS) is one of the largest open access only publishers. They
as of this writing have published 2.1 million articles. One of the strongs advantages of PLOS
is that they provide an API to their Solr instance, which is a very flexible way to
search their articles. The main R interace to PLOS is through [rplos][].

## Elsevier/Scopus

Elsevier is one of the largest publishers. Most of their articles are not open access. However,
they have a numbrer of advantages if you have access to their articles: they are one of the few
publishers to provide machine readable XML (many publishers do have XML versions of articles,
but do not provide it); they are one of the few (two) publishers part of Crossref's text
and data mining program. The packages [fulltext][] and [crminer][] can be used to access
Elsevier articles through Crossref's TDM program. There's an interface to Scopus article
search within [fulltext][].

## arXiv/bioRxiv

arXiv and bioRxiv are preprint publishers, the former in existence for many years, and the
latter new on the scene. You can access articles from these publishers through [fulltext][].
arXiv does provide a web API that we hook into; bioRxiv does not, but we can get you
articles nonetheless.

## Pubmed/PMC/Europe PMC

Pubmed/PMC is a corpus/website of NIH funded research in the United States; while Europe PMC
is an equivalent for the European Union. You can access articles from Pubmed/PMC through
[fulltext][], and for Europe PMC through [europepmc][].

## Microsoft Academic Research

Microsoft Academic Research (MAR) is a search engine for research articles. You can use their
GUI web interface to search, and they provide APIs for programmatic access. The R interface
for MAR is [microdemic][]; and [fulltext][] hooks into `microdemic` as well for article search
and abstract retrieval.

## Directory of Open Access Journals

Directory of Open Access Journals (DOAJ) maintains data on open access journals, as well
as some portion of the articles in those journals. Thus, you can search for journals as
well as articles with DOAJ. The R interface for DOAJ is [jaod][].

## JSTOR

JSTOR's Data for Research program gives institutions with access to JSTOR, access
to full text of articles within JSTOR. There is no way however to make the interaction
with JSTOR completely programmatic, thus making reproducible research very difficult.
Nonetheless, there is an R package ([jstor][]) for using data from JSTOR's Data
for Research.

## ORCID

ORCID (https://orcid.org/) is an organization keeping track of identifiers and metadata
for researchers around the world. Individuals can optionally maintain metadata on their
scholarly works connected to their account with ORCID. Thus, across all of ORCID, a
significant cache of metadata is accruing on scholarly works, their funding amounts,
collaborators, etc., useful for bibliometrics research and more. The R interface for
ORCID is [rorcid][].

## Citoid/Open Citation Corpus

The Open Citation Corpus (http://opencitations.net/) holds records of which articles
cite which other articles, allowing for all important research on the scholarly
web of citation. Citation data has been very closely guarded until recently, but the
largest publishers are still not contributing to the Open Citation Corpus. The R
interface to the Open Citation Corpus is [rcitoid][].

# fulltext: a swiss army knife for text mining in R

[fulltext][] is a general purpose R package for the data part of text-mining:
search for articles, get links to articles, get article abstracts, and fetch
full text of articles. The `fulltext` package is always adding additional
data sources as time allows (See Table 1). Starting from searching for
articles, the outputs of search can be fed into a function to get links
to those articles, or to get abstracts for those articles, or to fetch
their full text.

The following is a breakdown of the major distinct parts of `fulltext`.

## Search

`ft_search()` provides search access to nine different data sources (PLOS, BMC, Crossref,
Entrez, arXiv, bioRxiv, Europe PMC, Scopus, Microsoft Academic), creating a
mostly unified interface to all data sources. The parts of each data source
that are common are for the most part factored out into the parameters of the
`ft_search()` function: query term(s), pagination (number of results, result
number to start at). In addition, we allow the user to pass on data source specific
options to refine the search per data source.

With `ft_search()`, you can query any combination of the nine data sources at once.
The returned object is a list, with access to results of each data source by its name
(e.g., `$plos`, or `$crossref`). For each data source, the returned object does vary
because the returned data from each data source widely varies; for the most part
data.frame's are returned. For those data sources not queried, their slot is empty.

One important aspect of the research result we highlight is the licenses in the
returned data for each data source.

```{r}
x <- ft_search(query = 'ecology', from = c("plos", "crossref"))
```

The results for this PLOS search have all CC-BY licenses

```{r}
x$plos
```

Whereas the results for this Crossref search have mixed licenses

```{r}
x$crossref
```

You can dig into the license field for each article, with URLs holding information
on each license

```{r}
vapply(x$crossref$data$license, function(w) w$URL[1], "")
```

## Links

`ft_links()` provides two pathways to get links (URLs) for articles, with a choice
of four different data sources (PLOS, BMC, Crossref, Entrez). First, you can
use `ft_search()`, then pass the output of that function to `ft_links()`.

```{r}
out <- ft_search(query = "ecology", from = "entrez")
ft_links(out)
```

Second, you can pass DOIs directly to `ft_links()`. Both end up at the same point, links
for each article, if they could be found for the user selected data source.

```{r eval=FALSE}
# FIXME
ft_links(out$entrez$data$doi)
```

The biggest caveat with `ft_links()` is that we can't gaurantee that the links
will work. Link rot is one way in which the links may not work: link rot is when
the URL does not point to the original content anymore, or fails altogether.
Additionally, with Crossref, publishers can deposit URLs for articles, but they
make change the URLs at some later date but not update the URLs with Crossref.

## Abstracts

`ft_abstract()` provides access to article abstracts from four different
data sources (PLOS, Scopus, Microsoft Academic Research, Crossref). The only
way to use the function is to pass article identifiers, which are for the most
DOIs.

The advantage of abstracts over full text is that abstracts can often be
retrieved even for paywalled articles. That is, you can have much broader
coverage of the articles you're targeting relative to full text.

If you are after abstracts, and you are already getting or already have full text,
and if the articles are in XML format, then you can use [pubchunks][] to extract
out the abstracts.

## Fetch full text

`ft_get()` fetchs full text of articles from many different data sources. From the DOIs
that are passed in to the function, we detect the publisher, and there are specific plugins
for certain publishers: AAAS, American Institute of Physics, American Society of
Clinical Oncology, American Society for Microbiology, arXiv, bioRxiv, BiomedCentral,
Copernicus, Crossref, Elife, Elsevier, Pubmed/PMC via NCBI's Entrez, Frontiers, IEEE,
Informa, Instituto de Investigaciones Filologicas, American Medical Association,
Microbiology Society, PeerJ, Pensoft, PLOS, PNAS, Royal Society of Chemistry,
ScienceDirect, Scientific Societies, and Wiley.

If there's no built-in plugin for the publisher already, we use the FTDOI API
(<https://ftdoi.org>) to try to get the link for the full text of the article. If
the FTDOI API doesn't bear fruit, we search Crossref for a link to the full text.
If Crossref doesn't have any full text links, we give up.

Since users can go through a lot of article requests, we cache successfully
downloaded articles, and keep that knowledge consistent across R sessions; all
subsequent requests for the same article just use the cached version. Additionally,
all errors in `ft_get()` are collected in a tidy data.frame in the output of the
function to help the user quickly determine what went wrong.

# How to text mine from R: Three case studies

## Case study 1: Citation mining

```{r citation-mining, child="use-cases/citation-mining.Rmd", eval=TRUE}
```

## Case study 2: Abstract mining

```{r abstract-mining, child="use-cases/abstract-mining.Rmd", eval=TRUE}
```

## Case study 3: Full text mining

```{r full-text-mining, child="use-cases/full-text-mining.Rmd", eval=TRUE}
```

# Future directions

To make text mining easier:

1. publishers should provide XML if they have it
2. publishers should not change URL patterns so often, or at all
3. publishers should keep their Crossref metadata up to date
4. remove all paywalls (Easy, yes?)
5. ...

# Acknowledgments

xxxx

# Data Accessibility

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
<https://github.com/ropensci/textmine>. Software can be found at
<https://github.com/ropensci/xxx>, xxxx, all under MIT licenses.

# References

[rcrossref]: https://github.com/ropensci/rcrossref
[rorcid]: https://github.com/ropensci/rorcid
[fulltext]: https://github.com/ropensci/fulltext
[microdemic]: https://github.com/ropensci/microdemic
[arXiv]: https://github.com/ropensci/arXiv
[rdatacite]: https://github.com/ropensci/rdatacite
[rbhl]: https://github.com/ropensci/rbhl
[rentrez]: https://github.com/ropensci/rentrez
[rplos]: https://github.com/ropensci/rplos
[crminer]: https://github.com/ropensci/crminer
[europepmc]: https://github.com/ropensci/europepmc
[jaod]: https://github.com/ropensci/jaod
[jstor]: https://github.com/ropensci/jstor
[rcitoid]: https://github.com/ropenscilabs/rcitoid
[citecorp]: https://github.com/ropenscilabs/citecorp
[serrano]: https://github.com/sckott/serrano
[habanero]: https://github.com/sckott/habanero
[quanteda]: https://quanteda.io/
[pubchunks]: https://github.com/ropensci/pubchunks
[pdftools]: https://github.com/ropensci/pdftools
[libpoppler]: https://poppler.freedesktop.org/
[tesseract]: https://github.com/ropensci/tesseract