### 2019-08-23 use-cases/full-text-mining.Rmd

```{r include=FALSE}
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = '#>',
  collapse = TRUE,
  cache.path = "cache/"
)
```

In this example, xxxx

```{r}
library("fulltext")
# library("crminer")
```

_Search for articles_

Search for the term _ecology_ in PLOS journals.

```{r}
(res1 <- ft_search(query = 'ecology', from = 'plos'))
```

Each publisher/search-engine has a slot with metadata and data

```{r}
res1$plos
```

_Get full text_

Using the results from `ft_search()` we can grab full text of some articles

```{r}
(out <- ft_get(res1))
```

_Extract text from pdfs_

Ideally for text mining you have access to XML or other text based formats. However, 
sometimes you only have access to PDFs. In this case you want to extract text 
from PDFs. `fulltext` can help with that.

You can extract from any pdf from a file path, like:

```{r}
path <- system.file("examples", "example1.pdf", package = "fulltext")
ft_extract(path)
```

_Extract text chunks_

Requires the [pubchunks][] library. Here, we'll search for some PLOS articles, then
get their full text, then extract various parts of each article with `pub_chunks()`.

```{r}
library("pubchunks")
res <- ft_search(query = "ecology", from = "plos", limit = 3)
x <- ft_get(res)
x %>% ft_collect() %>% pub_chunks(c("doi", "history")) %>% pub_tabularize()
```

_Further processing_

From here, you'll likely want to use R packages for removing stopwords, tokenizing, 
and more. The Natural Language Processing CRAN Task View
(https://cloud.r-project.org/web/views/NaturalLanguageProcessing.html) is a good place
to start if you're not sure what packages to use. Some packages worth trying in this
space are [antiword][], [pdftools][], [rtika][], [tokenizers][], and [textreuse][].

[antiword]: https://github.com/ropensci/antiword
[pdftools]: https://github.com/ropensci/pdftools
[rtika]: https://github.com/ropensci/rtika
[tokenizers]: https://github.com/ropensci/tokenizers
[textreuse]: https://github.com/ropensci/textreuse